---
title: "Introduction to motiftools"
author: "Diego Diez"
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('motiftools')`"
output:
    rmarkdown::html_document
abstract: >
  Introduction to the motiftools package.
vignette: >
 %\VignetteIndexEntry{Introduction to motiftools}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
library(BiocStyle)
library(ggplot2)
library(ggthemes)
theme_set(theme_base(base_size = 12) + theme(plot.background = element_rect(color = NA)))
```

# Introduction

motiftools is a package for the analysis and interpretation of sequence motifs. Its initial implementation is focused on protein motif analysis obtained from the MEME suite, but its functionality may extend to other applications and tools, including analysis of nucleotide sequences and motif analysis tools found in Bioconductor or CRAN.

Motif ranges are stored as `RangedData` objects (`r Biocpkg("IRanges")`) package). This allows to include metadata, like p-value, score and other output from the search process.

```{r, message=FALSE}
library(motiftools)
library(motifStack)
```

# Plotting functions

## Alignments

You can align two protein sequences specifying a particular substitution matrix (e.g. BLOSUM62).

```{r}
data("BLOSUM62", package = "Biostrings")
align("EGGALAPA", "DGGIVLPSAV", score.matrix = BLOSUM62, type = "global")$alignment
align("EGGALAPA", "DGGIVFFRF", score.matrix = BLOSUM62, type = "local")$alignment
```

## Motifs

```{r}
# load MEME results.
f <- system.file("files/meme_ras/meme.xml", package = "motiftools")
x_meme <- readMEME(f)
m_meme <- getMotifMatrix(x_meme)
plotMotifMatrix(m_meme, fill = c("white", "grey"))

# load FIMO results.
f <- system.file("files/fimo_ras/fimo.xml", package = "motiftools")
x_fimo <- readFIMO(f)
m_fimo <- getMotifMatrix(x_fimo)
plotMotifMatrix(list(m_meme, m_fimo), fill = c("white", "grey"))

# load MAST results.
f <- system.file("files/mast_ras/mast.xml", package = "motiftools")
x_mast <- readMAST(f)
m_mast <- getMotifMatrix(x_mast)
plotMotifMatrix(list(m_meme, m_fimo, m_mast), fill = c("white", "grey"))

# highlight some sequences.
h <- grepl("RASH|RASK", sequenceNames(x_meme))
names(h) <- sequenceNames(x_meme)
plotMotifMatrix(list(m_meme, m_fimo, m_mast), fill = c("white", "grey"), high = h, high.col = c("black", "orange"))

# plot a motif model logo (from MEME output).
p <- pwm(x_meme)[["motif_4"]]
plotMotifLogo(p, col = rainbow(20))
```

You can compare the motifs obtained with MEME using the TOMTOM tool. Then you can read the results and look at the comparison matrix.

```{r}
f <- system.file("files/tomtom_ras/tomtom.xml", package = "motiftools")
x_tomtom <- readTOMTOM(f)
plotMotifMatches(x_tomtom)
plotMotifMatches(x_tomtom, fill = "e_value")
```


## Sequences

```{r}
m <- matrix(sample(1:5, 100*10, replace = TRUE), ncol = 100, nrow = 10, dimnames = list(letters[1:10], 1:100))
h <- hclust(dist(m))
plotConservationMatrix(m, h)
```

